---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---



```{r}
link='https://raw.githubusercontent.com/ksyanyu/543Final/main/deliverable3/PopulationGrowthAllCountry.csv'

library(rio)

#getting the data TABLE from the file in the cloud:
allCountry=rio::import(link)
```
```{r}

```

```{r}
# We are interested in the data of 2019
# keep the data of 2019 for all countries
all2019=allCountry
```

```{r}
all2019=all2019[all2019$Year==2019,]
```

```{r}
all2019[all2019==''] = NA
all2019=na.omit(all2019)
```

```{r}
# explore the behavior of the numerical variable
boxplot(all2019[,c('PopulationChange')])
```

```{r}
str(all2019)
```

```{r}
# transfer int into num
all2019[,4]=as.numeric(all2019[,4])
```




```{r}
str(all2019)
```

```{r}
all2019[all2019=='World'] = NA
all2019=na.omit(all2019)
```


```{r}
# cluster the data using **pam**

set.seed(123)

library(cluster)
var=c('PopulationChange', 'Year')

distMatrix=cluster::daisy(all2019[,var])
          
res.pam=cluster::pam(x=distMatrix,
                     k = 3,
                     cluster.only = F)

```


# save the cluster info into the data frame:

```{r}
all2019$cluster=as.factor(res.pam$clustering)
```

Check order:

```{r}
theVars=c('PopulationChange', 'Year', 'cluster')
aggregate(.~cluster,
          data=all2019[,theVars],
          FUN=median)
```

# recode so that number ordering make sense:

```{r}
all2019$cluster=factor(all2019$cluster,
                           levels=c(2,1,3),
                           labels=c("low","mid","high"), 
                           ordered=T)
```



```{r}
linkMap="https://github.com/EvansDataScience/VAforPM_Spatial/raw/main/worldMap.geojson" 

library(sf)
mapWorld=read_sf(linkMap)
```

This is the data that came with the map:

```{r}
# see data in map
head(mapWorld)
```

```{r}
# rename the column to make it easy to merge
names(all2019)=c("Country", "ISO3", "Year", "PopulationChange", "cluster")
```

```{r}
str(all2019)
```


# add our data to the map by merging:

```{r}
mapWorldVars=merge(mapWorld, #map first
                   all2019, 
                   by='ISO3') 
```

```{r}
str(mapWorldVars)
```

# 1. Base layer: 

```{r}
library(ggplot2)
# plot original map
base=ggplot(data=mapWorld) + geom_sf(fill='grey90',
                                     color=NA) + theme_classic()

base #base map

```

# 2. Data layer:

# This will be plot on top of the previous. Whatever is still in grey means a country has no information:

```{r}
colMap= base + geom_sf(data=mapWorldVars,
                       aes(fill=PopulationChange),
                       color=NA)
colMap
```

# alter the color sequence:

```{r}
colMap + scale_fill_gradient(low = 'yellow',
                             high= 'red')
```


# Now paint the countries using the cluster information (ggplot decides color):

```{r}
clusterMap= base + geom_sf(data=mapWorldVars,
                           aes(fill=cluster),
                           color=NA) # color of border
clusterMap 
```

# Now we choose the palette (get it from [here](http://colorbrewer2.org/)):

```{r}
clusterMap+ scale_fill_brewer(palette ='Blues')
```


# Some more customization:

```{r}
#previously

theLegTitle="Population growth\n(grey is missing)"

#now
clusterMap+ scale_fill_brewer(palette ='Blues',
                             
                              
                              name=theLegTitle)
```
